---
title: 'Supporting code C '
author: "Danika Hill"
date: "21/11/2019"
---

\pagebreak

This document contains code to use the Linear mixed effect regression to model cell type frequencies over time. 

This code can be used to recreate the following figures: 
  Fig 3C
  Fig 3D
  Fig 5B
  Fig 5C

Provided in the RData file is the Tanzanian data, pre-subsetted for ease of use : 
"Tanz.age" : dataset with 20 cells types identified as significantly different between B0 and B32 in paired analysis (% of parent population) 
"GCRF.GenR" : dataset with 19 cell types in common to GenR cohort, (cell type frequencies as % of parent gate) from 319 Tanzanian children for which data  was available for each cell type 

Raw data for the Dutch Generation R cohort is available upon collaboration request (https://generationr.nl/researchers/collaboration/)
Contact details: 
Prof dr. Henriette Moll h.a.moll@erasmusmc.nl
Prof.dr. Menno van Zelm menno.vanzelm@monash.edu

"GenR.common" file preprocessing: 
The absolute counts for Generation R data were transformed into "% of parent" frequencies (e.g. "CD4+ T cell", "CD8+ T cells" of "CD19+ B cell") in order to be comparable to the Tanzanian data. The 19 cell types that were comparably gated between the two cohorts were used. Refer to GCRF.GenR for these cell-type names. 
Convert GenR month of age to "ageAtVisit" variable in weeks. 

# load packages and data

```{r}
library(tidyverse)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(pheatmap)
library(MuMIn)
library(lme4)

## download the R data file - add to working directory
load(file = "data/Supp_Code_C_Lmer_modelling.RData")
```

## Lmer modelling of cell type frequencies in Tanzanian children - Fig 3C. 

```{r}
# The Tanz.age data-set contains the 20 cells types that were identified as significantly different between B0 and B32 in paired analysis. 
# This code genereates LMER models, which are used to predict a data set for 20-220 weeks, that is then used to generate the heatmap. 

model.results <- matrix(data = 0, ncol = 6, nrow = 20) #Initiate Matrix
colnames(model.results) <- c("int.Tanz","slopeTanz", "pvalue.age","pvalue.sex", "marginal.r2", "cond.R2")
row.names(model.results) <- colnames(Tanz.age[,1:20])  # name rows with cell_types

for (i in 1:20){
lmer_age <- lmer(Tanz.age[,i] ~ ageAtVisit + sex+ (1|PID), REML=FALSE, data=Tanz.age)
lmer_sex <- lmer(Tanz.age[,i] ~ sex+ (1|PID), REML=FALSE, data=Tanz.age)
nullmodel <- lmer(Tanz.age[,i] ~ (1|PID), REML=FALSE, data=Tanz.age)
model.results[i,1] <- (summary(lmer_age))$coefficients[1,1] #intercept Tanz
model.results[i,2] <- (summary(lmer_age))$coefficients[2,1] #slope Tanz
model.results[i,3] <-  (anova(nullmodel, lmer_sex, lmer_age))$Pr[3]# pvalue for age effects
model.results[i,4] <-  (anova(nullmodel, lmer_sex, lmer_age))$Pr[2]# pvalue for sex effects
model.results[i,5] <-  r.squaredGLMM(lmer_age)[1,1] #marginal r squared 
model.results[i,6] <-  r.squaredGLMM(lmer_age)[1,2] #conditional r squared 
}
model.results[,3] <- p.adjust(model.results[,3], method = "fdr")  # adjust p-values for multiple comparisons
model.results[,4] <- p.adjust(model.results[,4], method = "fdr")  # adjust p-values for multiple comparisons

model.results <- data.frame(model.results) # Note: use p-values and R^2 from model.results to generate the annotation boxes for Figure 3C. 

# use the slope and intersect data to predict the frequency of each cell type between week 20-220. 
Tanz.model = matrix(nrow=201, ncol = 20)  
for (i in 1:20){
for (p in 20:220){
Tanz.model[p-20,i] = model.results$int.Tanz[i]+ (model.results$slopeTanz[i]*p)
}}

# generate heatmap: 
plotting <- as.matrix(t(Tanz.model[,]))
rownames(plotting) <- colnames(Tanz.age[,1:20])
colnames(plotting) <- c(20:220)
plotting <- ((plotting[,] / plotting[,1]))-1  ###normalise to week 20 values
breaksList <- seq(from = -0.5, to = 3.5, by = 4/499)

  # pdf(file = "Tanz all heated.pdf", width=7, height=4)
p <- pheatmap(plotting, color = colorRampPalette(rev(brewer.pal(n = 11, name ="Spectral")))(500), kmeans_k = NA, breaks = breaksList, border_color = NA,
cellwidth = NA, cellheight = NA, scale = "none", cluster_rows = TRUE,
cluster_cols = FALSE, clustering_distance_rows = "euclidean",
clustering_distance_cols = F, clustering_method = "complete", cutree_rows = 3, cutree_cols =2,
legend = TRUE, legend_breaks = NA,
legend_labels = NA, annotation_row = NA, 
annotation_names_row = TRUE, annotation_names_col = TRUE,
drop_levels = TRUE, show_rownames = T, show_colnames = F, main = NA,
fontsize = 10, display_numbers = F, number_format = "%.2f", number_color = "grey30",
fontsize_number = 0.8 * fontsize, gaps_row = NULL, gaps_col = NULL,
labels_row = NULL, labels_col = NULL, filename = NA, width = NA,
height = NA, silent = FALSE)
  # dev.off()

```

## Lmer modelling of individual cell types in Tanzanian children - Fig 3D. 

```{r}
# specify population to graph. Eg: 
# use model-results for p-value and R^2
pop <- 10   #"43_Transitional.Bcells" 
pop <- 4   #" "32_CD27pos.Bmem"
pop <- 16 #"63_CD4.Tcm"  

pop.name <- names(Tanz.age)[pop]
pop.name

ggplot(data=Tanz.age, aes(x=ageAtVisit, y=(Tanz.age[,pop])))+
  geom_point(size=2.5, pch= 21, show.legend = F)+
  geom_abline(aes(intercept=model.results$int.Tanz[pop], slope=model.results$slopeTanz[pop]), colour="red", size=1.5)+
  coord_cartesian(xlim = c(20,220))+
  ylab(paste(pop.name,"(% of parent)", sep = " ")) + xlab("")+ 
  theme_classic()

```

## Tanzania vs Dutch children - Figure 5B : Generate Data-sets and perform modelling

Note: Raw data for GenR (Dutch) children is not provided. See code introduction for information about how to access this data.
The remaining code can be used to generate plots once raw data is obtained.

```{r}
#Subset data to include the following columns - make sure names match between GenR and GCRF data-sets
common <- c("IGD+27- Bnaive","B memory","IGD-27-","B IgMmem","CD27-IgG+","CD27-IgA+","IGD-27+","IgD-27+IgM+","CD27+IgG+","CD27+IgA+",
            "VD2+","CD4+ Tnaive","CD4+ Tcm","CD4+ TemROtotal","CD4+ TemRA total","CD8+Tnaive","CD8+Tcm","CD8+TemRO total","CD8+TemRA total", 
            "PID","ageAtVisit","sex","study")

########################################################################################################################
#
# INSERT GenR data at this stage, having obtained access and normalised it as described in the introduction to this .Rmd
#
# Downstream code expects an object called 'GenR.common'
#
# This .Rmd will not knit, as this chunk will not run, as 'GenR.common' does not exist on the github version.
#
#
########################################################################################################################


# join studies together
Two.studies.data <- rbind(GenR.common, GCRF.GenR)
Two.studies.data<- Two.studies.data[,common]

# include only data from  week 21-125
study.data <- Two.studies.data[Two.studies.data$ageAtVisit<125 & Two.studies.data$ageAtVisit>21,] #limit only to age range of interest

# perform lmer models
model.results <- matrix(data = 0, ncol = 9, nrow = 19)
model.results <- data.frame(model.results)
colnames(model.results) <- c("int.GCRF","int.GenR","slopeGCRF", "slopeGenR", "GCRF.SE", "GenR.SE", "pvalue.study", "pvalue.interaction", "pvalue.sex")
row.names(model.results) <- colnames(study.data[,1:19])

for (i in 1:19){
model0<-lmer(study.data[,i]~ageAtVisit+(1|PID), data=study.data, REML=FALSE)
model1<-lmer(study.data[,i]~ageAtVisit+sex+(1|PID), data=study.data, REML=FALSE)
model2<-lmer(study.data[,i]~ageAtVisit+study+sex+(1|PID), data=study.data, REML=FALSE)
model3<-lmer(study.data[,i]~ageAtVisit*study+sex+(1|PID), data=study.data, REML=FALSE)
anova(model0, model1, model2, model3)
model.results[i,1] <- (summary(model3))$coefficients[1,1] #intercept GCRF
model.results[i,2] <- (summary(model3))$coefficients[1,1] + (summary(model3))$coefficients[3,1]#intercept GenR 
model.results[i,3] <- (summary(model3))$coefficients[2,1] #slope GCRF
model.results[i,4] <- (summary(model3))$coefficients[2,1] +(summary(model3))$coefficients[5,1] #slope GenR
model.results[i,5] <- (summary(model3))$coefficients[1,2] #GCRF-SE
model.results[i,6] <-  (summary(model3))$coefficients[3,2] #GenR -SE
model.results[i,7] <-  (anova(model0, model1, model2, model3))$Pr[3]# pvalue for difference between studies
model.results[i,8] <-  (anova(model0, model1, model2, model3))$Pr[4]#pvalue for interactions (ageAtVisit * study) 
model.results[i,9] <- (anova(model0, model1, model2, model3))$Pr[2]#pvalue for sex
}
model.results$pvalue.study.fdr <- p.adjust(model.results$pvalue.study,"fdr")
model.results$pvalue.interaction.fdr <- p.adjust(model.results$pvalue.interaction,"fdr")
model.results$pvalue.sex.fdr <- p.adjust(model.results$pvalue.sex,"fdr")

# use the results of the lmer models to generate a predicted data set that spans 21-125 weeks of age
GCRF.model = matrix(nrow=105, ncol = 19)  
for (i in 1:19){
for (p in 21:125){
GCRF.model[p-20,i] = model.results$int.GCRF[i]+ (model.results$slopeGCRF[i]*p)
}}

GenR.model = matrix(nrow=105, ncol = 19)  
for (i in 1:19){
for (p in 21:125){
GenR.model[p-20,i] = model.results$int.GenR[i]+ (model.results$slopeGenR[i]*p)
}}

```

# Tanzania vs Dutch children - Figure 5B : generate heatmaps of T cells 

```{r}
breaksList <- seq(from = -1.0, to = 4.5, by = 5.5/499)  #use for heatmaps colour

#Tanzanian data first
plotting <- as.matrix(t(GCRF.model[,]))
plotting <- plotting[11:19,] # pull out only T cell subsets 
rownames(plotting) <- colnames(study.data[,11:19]) # pull out only T cell subsets names
colnames(plotting) <- c(21:125)
plotting.norm <- ((plotting[,] / plotting[,1]))-1 # normalise to Week 21

pheatmap(plotting.norm, color = colorRampPalette(rev(brewer.pal(n = 11, name ="Spectral")))(500),breaks = breaksList, 
         border_color = NA, cellwidth = NA, cellheight = NA, scale = "none", cluster_rows = F,cluster_cols = F , show_rownames = T, 
         show_colnames = F, fontsize = 10, display_numbers = F, number_format = "%.2f", number_color = "grey30",
         fontsize_number = 0.8 * fontsize, gaps_row = NULL, gaps_col = NULL,silent = FALSE)

#GenR data
plotting.GenR <- as.matrix(t(GenR.model[,11:19]))
rownames(plotting.GenR) <- colnames(study.data[,11:19])
colnames(plotting.GenR) <- c(21:125)
plotting.GenR <- ((plotting.GenR[,] / plotting[,1]))-1

#pdf(file = "New GenR heated T.pdf",width=6, height=6) #

pheatmap(plotting.GenR, color = colorRampPalette(rev(brewer.pal(n = 11, name ="Spectral")))(500),breaks = breaksList, 
         border_color = NA, cellwidth = NA, cellheight = NA, scale = "none", cluster_rows = F,cluster_cols = F , show_rownames = T, 
         show_colnames = F, fontsize = 10, display_numbers = F, number_format = "%.2f", number_color = "grey30",
         fontsize_number = 0.8 * fontsize, gaps_row = NULL, gaps_col = NULL,silent = FALSE)

#print out p-values for T cell pops in order
model.results[11:19,"pvalue.sex"]
model.results[11:19,"pvalue.study.fdr"]
model.results[11:19,"pvalue.interaction.fdr"]
rownames(model.results[11:19,])
```

# Tanzania vs Dutch children - Figure 5B  : generate heatmaps of B cells 

```{r}
breaksList <- seq(from = -1.0, to = 6.5, by = 7.5/499) #use for heatmaps colour

#Tanzanian data first
plotting <- as.matrix(t(GCRF.model[,1:10]))
rownames(plotting) <- colnames(study.data[,1:10])
colnames(plotting) <- c(21:125)
plotting.norm <- ((plotting[,] / plotting[,1]))-1

pheatmap(plotting.norm, color = colorRampPalette(rev(brewer.pal(n = 11, name ="Spectral")))(500),breaks = breaksList, 
         border_color = NA, cellwidth = NA, cellheight = NA, scale = "none", cluster_rows = F,cluster_cols = F , show_rownames = T, 
         show_colnames = F, fontsize = 10, display_numbers = F, number_format = "%.2f", number_color = "grey30",
         fontsize_number = 0.8 * fontsize, gaps_row = NULL, gaps_col = NULL,silent = FALSE)

#GenR data
plotting.GenR <- as.matrix(t(GenR.model[,1:10]))
rownames(plotting.GenR) <- colnames(study.data[,1:10])
colnames(plotting.GenR) <- c(21:125)
plotting.GenR <- plotting.GenR[1:10,]
plotting.GenR <- ((plotting.GenR[,] / plotting[,1]))-1

pheatmap(plotting.GenR, color = colorRampPalette(rev(brewer.pal(n = 11, name ="Spectral")))(500),breaks = breaksList, 
         border_color = NA, cellwidth = NA, cellheight = NA, scale = "none", cluster_rows = F,cluster_cols = F , show_rownames = T, 
         show_colnames = F, fontsize = 10, display_numbers = F, number_format = "%.2f", number_color = "grey30",
         fontsize_number = 0.8 * fontsize, gaps_row = NULL, gaps_col = NULL,silent = FALSE)

#print out p-values for B cell pops in order
model.results[1:10,"pvalue.sex"]
model.results[1:10,"pvalue.study.fdr"]
model.results[1:10,"pvalue.interaction.fdr"]

```
# Tanzanian vs Dutch - Figure 5C : Individual cell types
```{r}
#set theme
theme_set(theme_bw()+
  theme(axis.line = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", size = 1),
    panel.background = element_blank(), 
    axis.text=element_text(size=12), 
    axis.title=element_text(size=14), 
    plot.margin=grid::unit(c(10,5,0,5), "mm")))

pop <- 12 # change this to generate plots of different cell types

pop.name <- names(study.data)[pop]
pop.name

ggplot(data=study.data, aes(x=study.data$ageAtVisit, y=study.data[,pop], colour=study))+
  scale_colour_manual(values=c("black", "blue"))+
  geom_point(size=2.5, alpha = 0.4, show.legend = F)+
  geom_abline(aes(intercept=model.results$int.GCRF[pop], slope=model.results$slopeGCRF[pop]), colour="black", size=1.5)+
  geom_abline(aes(intercept=model.results$int.GenR[pop], slope=model.results$slopeGenR[pop]), colour="blue", size=1.5)+
  scale_x_continuous(expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0))+
  coord_cartesian(xlim = c(20,127), ylim = c(0,100))+ ## change the ylim for each cell type as necessary
  ylab(paste(pop.name,"(% of parent)", sep = " ")) + xlab("")

```

# Diffusion pseudotime analysis - Fig 5D-G
The method used here is based on the approach used in the following paper: 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6686855/
A clinically meaningful metric of immune age derived from high-dimensional longitudinal monitoring
Ayelet Alpert et al. Nat Med. 2019 Mar; 25(3): 487–495.  

```{r}
study.data <- Two.studies.data

# GenR
norm.study <- study.data[study.data$study %in% "GenR",] #select out only GenR data

conf.Int <- matrix(nrow = 19, ncol =4)
for(i in 1:19){
conf.Int[i,1] <- quantile(norm.study[,i], probs=c(0.1), na.rm= T) # 10th percentile
conf.Int[i,2] <- quantile(norm.study[,i], probs=c(0.9), na.rm= T) # 90th Percentile
conf.Int[i,3] <-  mean(norm.study[((norm.study[,i] > conf.Int[i,1]) & (norm.study[,i] < conf.Int[i,2])),i], na.rm= T) # determine mean of data within 90th percentile 
conf.Int[i,4] <-  sd(norm.study[((norm.study[,i] > conf.Int[i,1]) & (norm.study[,i] < conf.Int[i,2])),i], na.rm= T) # determine st dev of data within 90th percentile 
}
for(i in 1:19){
  norm.study[,i] <- (norm.study[,i] - conf.Int[i,3])/conf.Int[i,4] #normalise GenR data using this mean and stdev
}
GenR <- norm.study #write out new data-set

# apply mean and stdev from GenR to normalise GCRF data
norm.study <- study.data[study.data$study %in% "GCRF",]
for(i in 1:19){
  norm.study[,i] <- (norm.study[,i] - conf.Int[i,3])/conf.Int[i,4] #normalise GCRF data
}
GCRF <- norm.study

# Build one data-frame 
Both.normalised <- rbind(GenR,GCRF) 

# Perform PCA
study.pca <- prcomp(Both.normalised[,c(1:19)], center = TRUE,scale. = TRUE)
df_out <- as.data.frame(study.pca$x) #create dataframe of PC coprdinates
df_out$ageAtVisit <- norm.study$ageAtVisit # add covariates
df_out$group <- norm.study$study # add covariates
df_out$agec <- norm.study$agec # add covariates

# Plot PCA data:
ggplot(df_out,aes(x=PC1,y=PC2,color=df_out$ageAtVisit )) +geom_point() # PC1 = ageAtVisit

#keep only vars which correlate with PC1 = "AGE"
var <- get_pca_var(study.pca)
var$coord[,1:2]
which.vars <- as.data.frame(var$coord[,1:2])
which.vars <- which.vars[(which.vars$Dim.1 < c(-0.5)) |(which.vars$Dim.1 > c(0.5)) , ] # filter for cell types that correlate greater than 0.5 with PCR 
keep.vars <- row.names(which.vars) # 18 variables fally within this range and are used to create diffusion pseudotime. 

################# Diffusion map

GenR <- Both.normalised[Both.normalised$study %in% "GenR",]
GCRF.Bag <- Both.normalised[Both.normalised$study %in% "GCRF" & Both.normalised$site %in% "BAGAMOYO",]
study.data <- rbind(GenR, GCRF.Bag) # build post normalisation data-frame of 2222 samples
study.data$study <- as.factor(study.data$study)

dm <- DiffusionMap(study.data[,keep.vars]) # create diffusion map using 18 variables
dpt <- DPT(dm) # diffusion pseudotime
study.data$DC1 <- dm$DC1 #add coordinates to dataset
study.data$DC2 <- dm$DC2 #add coordinates to dataset
study.data$DC3 <- dm$DC3 #add coordinates to dataset

#calculate pseudotime and rescale 
pseudo <-  dpt@dm@sigmas@optimal_sigma #create pseudotime variable
range01 <- function(x){(x-min(x))/(max(x)-min(x))} #rescale function
pseudotime <- range01(pseudo)
study.data$pseudotime <- pseudotime #add normalised pseudotime to dataset

# evaluate all datapoints: Fig 5D
ggplot(study.data, aes(DC1, DC2, colour = pseudotime))+
  geom_point(alpha = 1, cex = 1)+
  theme_bw()+
  scale_color_viridis_c(begin = 1, end = 0.3)+
  coord_cartesian(xlim = c(-0.04, 0.0605), ylim = c(-0.05, 0.1005))+
  scale_x_continuous(expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0))+
  theme(panel.grid.minor = element_blank(),panel.grid.major = element_blank())

# evaluate the 20-125 wk time-point for each study:  Fig 5D
plotting <- study.data[study.data$ageAtVisit<125 & study.data$ageAtVisit>20,]

ggplot(plotting, aes(DC1, DC2, colour = pseudotime))+
  geom_point(alpha = 1, cex = 1)+
  theme_bw()+
  scale_color_viridis_c(begin = 1, end = 0.3)+
  coord_cartesian(xlim = c(-0.04, 0.0605), ylim = c(-0.05, 0.1005))+
  scale_x_continuous(expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0))+
  theme(panel.grid.minor = element_blank(),panel.grid.major = element_blank())+
  facet_grid(~study)
  
# plot cell types against diffusion map: Fig 5E
ggplot(study.data, aes(DC1, DC2, colour = study.data$`IGD+27- Bnaive`))+
  geom_point(alpha = 0.8, cex = 1)+
  theme_bw()+
  scale_color_viridis_c(option = "magma", begin = 1, end = 0.45)+
  coord_cartesian(xlim = c(-0.04, 0.0605), ylim = c(-0.05, 0.1005))+
  scale_x_continuous(expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0))+
  theme(panel.grid.minor = element_blank(),panel.grid.major = element_blank())

# plot Age against diffusion map: Supp Fig 11
ggplot(study.data, aes(DC1, DC2, colour = ageAtVisit))+
  geom_point(alpha = 1, cex = 1)+
  theme_bw()+
  scale_color_viridis_c(begin = 1, end = 0.3)+
  coord_cartesian(xlim = c(-0.04, 0.0605), ylim = c(-0.05, 0.1005))+
  scale_x_continuous(expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0))+
  theme(panel.grid.minor = element_blank(),panel.grid.major = element_blank())+
  facet_grid(~study)

# violin plots for pseudotime vs study  Fig 5F
ggplot(plotting, aes(study, pseudotime, col = study))+
  geom_violin(scale = "width") +
  geom_jitter(width = 0.1, alpha = 0.75,cex = 2, pch = 21) +
  scale_colour_manual(values = c("black", "blue"))+
  stat_summary(fun.y = "median", geom = "point")+
  theme_classic()+
  coord_cartesian(ylim = c(0, 0.5))+
  scale_y_continuous(expand = c(0, 0))

wilcox.test(plotting[plotting$study %in% "GCRF",]$pseudotime, 
            plotting[plotting$study %in% "GenR",]$pseudotime, paired = F) # determine p-value

# violin plots for Age vs study  Fig 5G
ggplot(plotting, aes(study, ageAtVisit, col = study))+
  geom_violin(scale = "width") +
  geom_jitter(width = 0.1, alpha = 0.75,cex = 2, pch = 21) +
  scale_colour_manual(values = c("black", "blue"))+
  stat_summary(fun.y = "median", geom = "point")+
  theme_classic()
  coord_cartesian(ylim = c(0, 0.5))+
  scale_y_continuous(expand = c(0, 0))

  wilcox.test(plotting[plotting$study %in% "GCRF",]$ageAtVisit, 
              plotting[plotting$study %in% "GenR",]$ageAtVisit, paired = F) # determine p-value
  
# determine which cell subsets correlate with pseudotime
correlation <- matrix(data = 0, ncol = 2, nrow = 19)
for (i in 1:19){
correlation[i,1] <-cor.test(x = study.data[,i], y = study.data[,"pseudotime"], 
                            method = c("spearman"))$estimate
correlation[i,2] <-cor.test(x = study.data[,i], y = study.data[,"pseudotime"], 
                            method = c("spearman"))$p.value
}
rownames(correlation) <- colnames(study.data[,1:19])
#plot as a heatmap
breaksList <- seq(from = -1.0, to = 1, by = 2/20)
correlation <- correlation[rev(order(correlation[,1])),]

pheatmap(correlation, color = colorRampPalette(rev(brewer.pal(n = 10, name ="Spectral")))(20),
         cluster_cols = F, cluster_rows = F, breaks = breaksList)
```